---
title: "Anwendung Dijkstra"
author: "Martin Zaefferer"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Anwendung Dijkstra

```{r}
require(GGally)
```


## Implementation Dijkstra in R

Hilfsfunktion zum erstellen des Lösungspfades als Sequenz von Knoten:
```{r}
getpath <- function(previous,start,end){
  #' @param previous vector, where each integer value indicates the node preceding another node in the graph
  #' @param start start node of the path
  #' @param start end node of the path
  
  i = end
  ## optionally, generate an adjacency matrix for the solution:
  #msol <- matrix(0,nrow=nrow(m),ncol=ncol(m)) 
  path <- end
  while(i != 0){
    ii <- previous[i]
    #msol[ii,i] <- 1
    i=ii
    if(i!= 0)
      path <- c(ii,path)
  }
  #msol
  path
}
```

Hauptfunktion:
```{r}
## Based on, with changes:
## https://www.algorithms-and-technologies.com/dijkstra/r
## ClaasM
## Posted: 2019-12-21 13:14, Last Updated: 2019-12-21 13:14
dijkstra <- function(graph, start, end=FALSE){
  #' Implementation of dijkstra using adjacency matrix.
  #' @param graph an adjacency-matrix-representation of the graph where (x,y) is the weight of the edge or 0 if there is no edge.
  #' @param start the node to start from.
  #' @param end the node to end at, optional. 
  #' Compute all shortest paths if not specified. 
  #' Set to FALSE to deactivate this stopping criterion (default).
  #' @return an array containing the shortest distances from the given start node to each other node
  
  
  # This contains the distances from the start node to all other nodes
  distances = rep(Inf, nrow(graph))
  
  # This contains whether a node was already visited
  visited = rep(FALSE, nrow(graph))
  
  # This contains the previous node of a current node in the path
  previous = rep(NA, nrow(graph))
  
  # The distance from the start node to itself is of course 0
  distances[start] = 0
  
  # And it has no previous node
  previous[start] = 0
  
  # And it becomes the current node
  current_node_index <- start
  
  # While there are nodes left to visit...
  while(any(!visited)){
    
    # Current node is visited
    visited[current_node_index] <- TRUE 
    
    # Stop if this is the end node
    if(current_node_index==end){
      break;
    }
    
    # ... then, for all neighboring nodes that haven't been visited yet....
    for(i in seq_along(graph[current_node_index,])) {
      # ...if the path over this edge is shorter...
      if(graph[current_node_index,i] != 0 && !visited[i] && distances[i] > distances[current_node_index] + graph[current_node_index,i]){
        # ...save this path as new shortest path.
        distances[i] = distances[current_node_index] + graph[current_node_index,i]
        previous[i] = current_node_index
      }
    }
    
    # ... find the node with the currently shortest distance from the start node...
    shortest_distance = Inf
    current_node_index = -1
    for(i in seq_along(distances)) {
      # ... by going through all nodes that haven't been visited yet
      if(distances[i] < shortest_distance && !visited[i]){
        shortest_distance = distances[i]
        current_node_index = i
      }
    }
  }
  
  ## If no end node: determine ALL shortest path from start
  if(!end){
    path <- NULL
    for(i in 1:nrow(m)){
      path <- getpath(previous,start,i)
    }
  }else{ ## Else, determine only shortest path from start to end node
    path <- getpath(previous,start,end)
  }
  
  
  list(distances=distances,previous=previous,path=path)
}
```

Anwendung des Algorithmus auf das Vorlesungsbeispiel:

```{r}
## original graph for the problem
m <- matrix(
  c(0, 1, 0, 0, 4, 0,
    1, 0, 1, 0, 1, 0,
    0, 1, 0, 1, 0, 0,
    0, 0, 1, 0, 6, 1,
    4, 1, 0, 6, 0, 0,
    0, 0, 0, 1, 0, 0),
  nrow = 6, ncol = 6
)
start = 1 ## start node
end = 6 ## end node
#res <- dijkstra(m,start=start) ## solver without end node
res <- dijkstra(m,start=start,end=end) ## solver with end node
res

## plotting the solution
require(network)
require(ggnet)
rownames(m) = 1:nrow(m)
colnames(m) = 1:nrow(m)
cols <- rep("black",nrow(m))
pathnodes <- (1:nrow(m)) %in% res$path
cols[pathnodes] <- "red"
net <- network(m, directed=F, names.eval = "weights", ignore.eval = FALSE)
set.seed(43)
ggnet2(net, 
       mode = "fruchtermanreingold", size = 10, label = T, 
       vjust = 0.5,
       node.color = cols, 
       label.color = "grey80", arrow.size = 10, edge.label = "weights", 
       arrow.gap = 0.05)
```



## Aufgabe

- Teil 1: Wenden Sie den Algorithmus auf das Ravensburg-Immenstaad Beispiel aus der Vorlesung an.

- Teil 2: In der Vorlesung wurde bereits die Frage gestellt: 
    
    - "Wie könnte man mit Dijkstra effizient nach allen den kürzesten Pfaden zwischen allen Knoten suchen (d.h., für beliebige Start- und Zielknoten)?"
    
    - Setzen Sie eine Antwort in R um, d.h., als Wrapper für die Dijkstra Implementierung:
    
```{r,eval=FALSE}
allpaths <- function(graph){
  
  ...
  
  dijkstra(graph= ...,start = ...) 
  
  ...

}
allpaths(m)
```
    
